---
title: "Data Analysis: Airport Traffic, Monte Carlo Numerical Integration, and Generative AI"
author: "Daniel Carney"
date: "2026-04-16"
format:
  pdf:
    toc: true
    toc-depth: 3
    number-sections: true
    colorlinks: true
    highlight-style: tango
    documentclass: article
    papersize: letter
    margin-top: 1in
    margin-bottom: 1in
    margin-left: 1in
    margin-right: 1in
    keep-tex: false
---

```{r}
#| label: setup
#| include: false
#| warning: false
#| message: false

# ---- Load Packages ----
library(tidyverse)
library(rvest)
library(janitor)
library(knitr)
library(kableExtra)
library(ggplot2)
```

This document presents four sections: the first two sections reference work 
completed in previous homework assignments,
examining global airport traffic patterns and demonstrating Monte Carlo numerical
integration. The third section explores the planning process for using generative
AI tools in data analysis, and the final section offers a reflection on concepts
learned throughout the course. All underlying code is contained in the Code
Appendix at the end of this document.

# Busiest Airports Analysis

## Data Collection and Preparation

The data for this analysis was scraped from Wikipedia's "List of Busiest
Airports by Passenger Traffic" page, which publishes annual rankings of the
world's busiest airports based on total passenger amounts Using the `rvest` package in R, six consecutive annual tables (2020 through 2025) were programmatically scraped from the webpage.
This used the `html_elements()` and `html_table()` functions to extract tabular 
data directly from the HTML source.

Six airports were selected for focused analysis: Hartsfield-Jackson Atlanta
International (ATL), Dallas/Fort Worth International (DFW), Denver
International (DEN), Los Angeles Inter

After scraping, the data required cleaning steps. Column names were
standardized using the `clean_names()` function from the `janitor` package to
ensure consistency across all six annual tables. The `totalpassengers` column
contained reference annotations in brackets (e.g., `[1]`) that were stripped
using a regular expression, after which the values were converted from
character strings to numeric. Missing values in the `rankchange` column,
which appeared when an airport's rank remained unchanged from the prior year,
were replaced with zero. Finally, the six annual tables were combined into a
single data frame and filtered to retain only the six airports of interest.

## Passenger Traffic Table

```{r}
#| label: tbl-airports
#| tbl-cap: "Rank and Total Passengers of Six Major International Airports (2020--2025)"
#| echo: false
#| warning: false
#| message: false

# ---- Step 1: Scrape Data from Wikipedia ----
airportRawList <- read_html(
  x = "https://en.wikipedia.org/wiki/List_of_busiest_airports_by_passenger_traffic"
) |>
  html_elements(css = "table") |>
  html_table()

# ---- Step 2: Extract and Label Each Year's Table ----
year_2025 <- airportRawList[[2]] |> clean_names() |> mutate(year = 2025)
year_2024 <- airportRawList[[3]] |> clean_names() |> mutate(year = 2024)
year_2023 <- airportRawList[[4]] |> clean_names() |> mutate(year = 2023)
year_2022 <- airportRawList[[5]] |> clean_names() |> mutate(year = 2022)
year_2021 <- airportRawList[[6]] |> clean_names() |> mutate(year = 2021)
year_2020 <- airportRawList[[7]] |> clean_names() |> mutate(year = 2020)

# ---- Step 3: Combine All Years ----
all_airports <- bind_rows(year_2025, year_2024, year_2023,
                          year_2022, year_2021, year_2020)

# ---- Step 4: Clean and Filter ----
airports_clean <- all_airports |>
  mutate(
    totalpassengers = str_remove_all(totalpassengers, "[^0-9]"),
    totalpassengers = as.numeric(totalpassengers),
    rankchange = replace_na(rankchange, 0)
  ) |>
  select(
    year, rank, code_iata_icao, airport, location,
    country, totalpassengers, rankchange
  ) |>
  filter(code_iata_icao %in% c(
    "ATL/KATL", "FRA/EDDF", "DXB/OMDB",
    "LAX/KLAX", "DFW/KDFW", "DEN/KDEN"
  ))

# ---- Step 5: Build the Table ----
airports_clean |>
  arrange(year, rank) |>
  mutate(totalpassengers = format(totalpassengers, big.mark = ",")) |>
  kable(
    col.names = c(
      "Year", "Rank", "Code", "Airport", "Location",
      "Country", "Total Passengers", "Rank Change"
    ),
    align = "c"
  ) |>
  kable_styling(
    latex_options = c("striped", "hold_position", "scale_down"),
    full_width = FALSE
  )
```

@tbl-airports presents the rank and total passenger counts for each of the
selected airports from 2020 through 2025. Several notable patterns emerge from
the data. Atlanta (ATL) consistently holds the top rank across every year in
the dataset. The year 2020 stands out because every
airport experienced a sharp decline in passenger traffic. 

## Passenger Traffic Visualization

```{r}
#| label: fig-airports-plot
#| fig-cap: "Total Passengers vs. Rank for Six Major International Airports (2020--2025)"
#| echo: false
#| warning: false
#| message: false
#| fig-width: 9
#| fig-height: 6

ggplot(airports_clean, aes(
  x = totalpassengers,
  y = rank,
  color = code_iata_icao,
  group = code_iata_icao,
  shape = factor(year)
)) +
  geom_line(linewidth = 1) +
  geom_point(size = 4) +
  scale_y_reverse() +
  scale_color_manual(
    values = c(
      "ATL/KATL"  = "#0072B2",
      "DFW/KDFW"  = "#D55E00",
      "DEN/KDEN"  = "#009E73",
      "LAX/KLAX"  = "#E69F00",
      "FRA/EDDF"  = "#CC79A7",
      "DXB/OMDB"  = "#56B4E9"
    ),
    labels = c(
      "ATL/KATL"  = "ATL",
      "DFW/KDFW"  = "DFW",
      "DEN/KDEN"  = "DEN",
      "LAX/KLAX"  = "LAX",
      "FRA/EDDF"  = "FRA",
      "DXB/OMDB"  = "DXB"
    )
  ) +
  scale_shape_manual(
    values = c(16, 17, 15, 18, 3, 7),
    name = "Year"
  ) +
  labs(
    x = "Total Passengers",
    y = "Global Rank",
    color = "Airport",
    shape = "Year"
  ) +
  theme_classic() +
  theme(
    text = element_text(size = 12),
    legend.position = "right",
    plot.title = element_text(face = "bold")
  )
```

@fig-airports-plot visualizes the relationship between total
annual passengers and global ranking position for each airport, with each
observation year distinguished by a unique shape. This plot reveals the
relationship between passenger volume and rank: airports with higher
passenger counts occupy the numerically lower (better) rank positions at the
top of the vertical axis. The inverted y-axis was chosen so that
the most highly ranked airports appear at the top of the display.

Atlanta's trends are clear as a stand-out cluster in the
upper-right corner of the plot. The 2020 data points for every airport are clustered toward the
left side of the horizontal axis. As the years progress from 2021
through 2025, the points migrate rightward, illustrating the gradual growth.
DXB's line, for example, follows a
visibly different trajectory from the North American airports, confirming the
observation from the table that Dubai's growth was faster and
more pronounced. 

# Monte Carlo Numerical Integration

## The Weibull Distribution Function

The function chosen for numerical integration via Monte Carlo simulation is the
Weibull probability density function (PDF) with a shape parameter of 1.5 and a
scale parameter of 1.

where $k = 1.5$ is the shape parameter and $\lambda = 1$ is the scale
parameter. With these parameter values, with its peak occurring at 
approximately $x \approx 0.63$ and
the bulk of its probability mass concentrated between $x = 0$ and $x = 2$.
The function is defined for all $x \geq 0$ and approaches zero
as $x$ increases. Because the Weibull PDF integrates to exactly 1 over its
entire support , the expected
value of the integral from 0 to 4 should be very close to 1, since nearly all
of the distribution's mass falls within this interval.

## Monte Carlo Methodology

Monte Carlo numerical integration is a technique that estimates
the value of a definite integral by exploiting random sampling.If we draw a
large number of random points uniformly within a bounding rectangle around a
region of interest, the proportion of those points that fall below the curve 
provide an estimate of the ratio of the area under the curve to the total area 
of the rectangle. By multiplying this proportion by the rectangle's area, we obtain an estimate of
the integral.

In this analysis, the bounding rectangle extends from $x = 0$ to $x = 4$ on
the horizontal axis and from $y = 0$ to $y = 0.8$ on the vertical axis,
yielding a total rectangle area of 3.2. For each random point $(x_i, y_i)$
drawn within this rectangle, we evaluate $f(x_i)$ and determine whether
$y_i \leq f(x_i)$. Points satisfying this condition lie within the area under
the curve and are counted as "inside," while those above the curve are counted
as "outside." 



$A_{\text{rect}} = 3.2$ is the area of the bounding rectangle,
$n_{\text{inside}}$ is the number of points falling under the curve, and
$n_{\text{total}}$ is the total number of sampled points. The accuracy of this
estimate improves as $n_{\text{total}}$ increases, governed by the Central Limit
Theorem, which guarantees that the sampling error decreases at a rate
proportional to $1/\sqrt{n}$.

## Results and Visualization

```{r}
#| label: fig-monte-carlo
#| fig-cap: "Monte Carlo Integration of the Weibull PDF (shape=1.5, scale=1) at Four Simulation Resolutions"
#| fig-subcap: ["n = 100 points", "n = 500 points", "n = 2{,}000 points", "n = 10{,}000 points"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 7

# ---- Define the Weibull PDF ----
f <- function(x) {
  dweibull(x, shape = 1.5, scale = 1)
}

# ---- Monte Carlo Simulation Function ----
# Returns a data frame of sampled points and the integral estimate.
monte_carlo_sim <- function(n, func, x_min, x_max, y_min, y_max) {
  x_vals <- runif(n, min = x_min, max = x_max)
  y_vals <- runif(n, min = y_min, max = y_max)
  f_x    <- func(x_vals)

  df <- data.frame(
    x    = x_vals,
    y    = y_vals,
    f_x  = f_x
  ) |>
    mutate(flag = ifelse(y <= f_x, "inside", "outside"))

  area       <- (x_max - x_min) * (y_max - y_min)
  proportion <- mean(y_vals <= f_x)
  estimate   <- area * proportion

  return(list(df = df, estimate = estimate))
}

# ---- Helper to Build a Plot ----
build_mc_plot <- function(result, n_label) {
  ggplot(result$df, aes(x = x, y = y)) +
    geom_point(aes(color = flag), alpha = 0.4, size = 0.8) +
    stat_function(
      fun = dweibull,
      args = list(shape = 1.5, scale = 1),
      xlim = c(0, 4),
      color = "black",
      linewidth = 1
    ) +
    scale_color_manual(
      values = c("inside" = "#0072B2", "outside" = "#D55E00")
    ) +
    labs(
      x = "x",
      y = "f(x)",
      color = "Point Location",
      title = paste0("Monte Carlo Integration (", n_label, ")"),
      subtitle = paste("Estimate =", round(result$estimate, 4))
    ) +
    theme_minimal(base_size = 10) +
    theme(legend.position = "bottom")
}

# ---- Run Simulations at Four Resolutions ----
set.seed(42)

res_100    <- monte_carlo_sim(100,    f, 0, 4, 0, 0.8)
res_500    <- monte_carlo_sim(500,    f, 0, 4, 0, 0.8)
res_2000   <- monte_carlo_sim(2000,   f, 0, 4, 0, 0.8)
res_10000  <- monte_carlo_sim(10000,  f, 0, 4, 0, 0.8)

# ---- Display the Four-Pane Small Multiple ----
build_mc_plot(res_100,   "n = 100")
build_mc_plot(res_500,   "n = 500")
build_mc_plot(res_2000,  "n = 2,000")
build_mc_plot(res_10000, "n = 10,000")
```

@fig-monte-carlo presents a small multiple visualization of
the Monte Carlo integration process at four simulation resolutions: 100, 500,
2,000, and 10,000 randomly sampled points. Each sub-panel displays the
Weibull PDF as a smooth black curve, overlaid with the randomly sampled
points colored blue if they fall inside and orange if they fall outside (above the curve). 
The subtitle of each panel reports the
corresponding integral estimate.

Several observations emerge from this visualization. At the lowest resolution
($n = 100$), the estimate is not accurate because of the lower sampling.
As the number of points increases to 500, the estimate improves, 
At $n = 2,000$, the estimate becomes accurate, and resembles the shape of the Weibull PDF.
Finally, at $n = 10,000$, the points produce an estimate
that is very close to the true value.


# Planning and Prompting GenAI Tools
## Data Context and Tidying Plan

## Data Context and Tidying Plan

The data for this section come from a study that followed 31 women
over the course of four years. At the start of the study, researchers measured
the amount of calcium in the ulna. Fifteen of the women received no special treatment and comprise
the null treatment (control) group, while the remaining sixteen women received
dietary support and followed a regular exercise program, comprising the
diet/exercise treatment group. The raw dataset, `calcium.csv`, is structured
with eight columns: the first four columns hold the control group's
measurements (Initial, Year 1, Year 2, Year 3) and the last four columns hold
the treatment group's measurements at the same time points. This format is
not tidy.

**Goal**: Tidy the calcium table into a glyph-ready format and then create a
line chart that shows the difference in ulnar calcium measurements over time
between the treatment group and the null group.

**Needs**:

*Nouns*: RStudio, calcium data, a blank R script, calcium cleaned data, a
line chart.

*Verbs*: Open RStudio, create a blank R script, load the data, clean the data,
pivot the data, create a new column to identify treatment or null group, create
a visualization.

**Steps**:

1. Open RStudio.
2. Create a blank R script and load the data from the calcium table.
3. Load any necessary packages for cleaning and creating visualizations
   (`tidyverse` and `ggplot2`).
4. Separate the two groups and create a new column assigning an ID to each
   row.
5. Use `pivot_longer()` to ensure that each row contains a single subject's
   measurement at a single time point.
6. Using `ggplot2`, build a line chart with time (Initial, Year 1, Year 2,
   Year 3) on the x-axis and the calcium values on the y-axis.
7. Assign a distinct color to the treatment group and a different color to the
   null group.
8. Create labels and a title for the visualization.
9. Run the code and ensure the output is as intended.
10. Draw conclusions regarding the effectiveness of the treatments based on
    the visualization.

## Plan Informed-Response

In this subsection, the calcium dataset and the tidying plan developed above
were prompted to Claude alongside the original data file. Claude's full response,
including its analysis of the data and the generated R code, is presented
below.

**Tool Used**: Claude

**Prompt**:

```
I have the attached data set which involves data drawn from 31 women over
the course of four years, the amount of calcium in the ulna for each woman
was measured, 15 did not change anything and the other 16 were in a treatment
group. I have created this plan for tidying the data and creating a
visualization, help me create the R code. Goal: Tidy the calcium table to a
glyph ready form and then create a line chart which shows the difference in
ulnar calcium measurements over time between the treatment group and the null
group.

Needs:
Nouns: Rstudio, Calcium data, Blank Rscript, calcium cleaned data, a line
chart
Verbs: Open Rstudio, create a blank rscript, load the data, clean the data,
pivot the data, create new column to identify treatment or null, create a
visualization

Steps:
1: Open RStudio
2: Create a blank RScript and load the data from the calcium table
3: Load any necessary packages for cleaning and creating visualizations
   (tidyverse and ggplot2)
4: Separate the two groups and create a new column and assigning an "ID" to
   each row
5: Use pivot_longer to ensure that each row contains a single subject
6: Using ggplot build a line chart with time (initial, year 1, year 2,
   year3) on the x-axis and the calcium values on the Y axis.
9: Assign a color to the treatment group and then a different color to the
   null group.
10: Create labels and a title for the visualization
11: Run the code and ensure the output is as intended.
12: Draw conclusions regarding the effectiveness of the treatments based on
    the visualization.
```

**Response**:

```
I can see the data clearly. The left 4 columns (y0-y3) are the control group
(15 women, with a missing last row), and the right 4 columns (y0_1-y3_1) are
the treatment group (16 women). Here's a walkthrough of the key decisions
made in the script:

Data entry & structure. The image shows the control group in the left 4
columns (15 rows) and the treatment group in the right 4 columns (16 rows,
with the asterisk row in the control side being blank). The data is entered
as two separate data frames reflecting this split.

Tidying with pivot_longer. Each row becomes one observation per subject per
time point --- so a subject with 4 measurements becomes 4 rows. The raw
column names (y0-y3) are recoded to readable labels (Year 0-Year 3) and
stored as an ordered factor so ggplot respects the correct time sequence on
the x-axis.

Two-layer visualization. The chart draws individual subject lines at low
opacity (so you can see variation within each group) with bold group mean
lines on top. This is more informative than just plotting means alone.

Conclusions to look for once you run it:

If the Treatment line stays higher or declines less than the Control line
from Year 0 -> Year 3, the treatment helped preserve ulnar calcium.
Check that Year 0 baselines are similar between groups --- if they're not,
it complicates interpretation.
A sharp Year 3 dip in both groups would suggest a time-based effect beyond
any treatment benefit.

# ============================================================
# Ulnar Calcium Study: Data Tidying & Visualization
# 31 women measured over 4 years (Year 0-3)
# Control group (n=15): y0, y1, y2, y3
# Treatment group (n=16): y0_1, y1_1, y2_1, y3_1
# ============================================================

# --- Step 1: Load packages ---
library(tidyverse)
library(ggplot2)

# --- Step 2: Enter the raw data ---

# Control group (15 women, no treatment)
control_raw <- data.frame(
  y0 = c(87.3, 59.0, 76.7, 70.6, 54.9, 78.2, 73.7, 61.8, 85.3, 82.3,
         68.6, 67.8, 66.2, 81.0, 72.3),
  y1 = c(86.9, 60.2, 76.5, 76.1, 55.1, 75.3, 70.8, 68.7, 84.4, 86.9,
         65.4, 69.2, 67.0, 82.3, 74.6),
  y2 = c(86.7, 60.0, 75.7, 72.1, 57.2, 69.1, 71.8, 68.2, 79.2, 79.4,
         72.3, 66.3, 67.0, 86.8, 75.3),
  y3 = c(75.5, 53.6, 69.5, 65.3, 49.0, 67.6, 74.6, 57.4, 67.0, 77.4,
         60.8, 57.9, 56.2, 73.9, 66.1)
)

# Treatment group (16 women)
treatment_raw <- data.frame(
  y0 = c(83.8, 65.3, 81.2, 75.4, 55.3, 70.3, 76.5, 66.0, 76.7, 77.2,
         67.3, 50.3, 57.7, 74.3, 74.0, 57.3),
  y1 = c(85.5, 66.9, 79.5, 76.7, 58.3, 72.3, 79.9, 70.9, 79.0, 74.0,
         70.7, 51.4, 57.0, 77.7, 74.7, 56.0),
  y2 = c(86.2, 67.0, 84.5, 74.3, 59.1, 70.6, 80.4, 70.3, 76.9, 77.8,
         68.9, 53.6, 57.5, 72.6, 74.5, 64.7),
  y3 = c(81.2, 60.6, 75.2, 66.7, 54.2, 68.6, 71.6, 64.1, 70.3, 67.9,
         65.9, 48.0, 51.5, 68.0, 65.7, 53.0)
)

# --- Step 3: Add group labels and subject IDs ---
control_raw <- control_raw %>%
  mutate(
    ID    = paste0("C", row_number()),
    group = "Control"
  )

treatment_raw <- treatment_raw %>%
  mutate(
    ID    = paste0("T", row_number()),
    group = "Treatment"
  )

# --- Step 4: Combine and pivot to long (tidy) format ---
calcium_long <- bind_rows(control_raw, treatment_raw) %>%
  pivot_longer(
    cols      = c(y0, y1, y2, y3),
    names_to  = "time_raw",
    values_to = "calcium"
  ) %>%
  mutate(
    time = recode(time_raw,
                  "y0" = "Year 0",
                  "y1" = "Year 1",
                  "y2" = "Year 2",
                  "y3" = "Year 3"),
    time = factor(time, levels = c("Year 0", "Year 1", "Year 2",
                                   "Year 3"))
  )

# --- Step 5: Compute group means per time point ---
calcium_summary <- calcium_long %>%
  group_by(group, time) %>%
  summarise(mean_calcium = mean(calcium), .groups = "drop")

# --- Step 6: Build the visualization ---
ggplot() +
  geom_line(
    data    = calcium_long,
    mapping = aes(x = time, y = calcium, group = ID, color = group),
    alpha   = 0.20,
    linewidth = 0.4
  ) +
  geom_line(
    data    = calcium_summary,
    mapping = aes(x = time, y = mean_calcium, group = group, color = group),
    linewidth = 1.4
  ) +
  geom_point(
    data    = calcium_summary,
    mapping = aes(x = time, y = mean_calcium, color = group),
    size    = 3
  ) +
  scale_color_manual(
    values = c("Control" = "#E07A5F", "Treatment" = "#3D9B8F"),
    name   = "Group"
  ) +
  labs(
    title    = "Ulnar Calcium Measurements Over Time",
    subtitle = "Bold lines show group means; faint lines show individual subjects",
    x        = "Time Point",
    y        = "Ulnar Calcium (mg)",
    caption  = "Control n=15 | Treatment n=16"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title    = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(color = "gray40", size = 10),
    legend.position = "top",
    panel.grid.minor = element_blank()
  )

ggsave("calcium_lineplot.png", width = 8, height = 5.5, dpi = 150)
```

## Generic Response

In contrast to the plan-informed approach above, this subsection presents the
results obtained when the same AI tool was given a generic, unstructured
prompt alongside the same calcium dataset. 

**Tool Used**: Claude 

**Prompt**:

```
Here is a data set that I need you to use R to clean and make a visualization
showing the trends of the responses over time
```

**Response**:

```
Looking at your data, I can see 8 variables (y0-y3 and y0_1-y3_1) across 14
rows, with some missing values (*) in the last row. Here's clean,
well-commented R code to wrangle and visualize it:

Here's what the script does, step by step:

Data wrangling:

All values are entered manually (since the * cells are missing, they're
treated as NA and dropped with filter(!is.na(value)))
A time index is added per row, then the data is pivoted from wide to long
format
Two derived columns are created: group (Group 1 = y0-y3, Group 2 = y0_1-y3_1)
and timepoint (the 0-3 suffix, representing the within-group time progression)

Visualization:

A summary table of mean +/- SE is computed per group x timepoint
The plot layers: raw jittered points -> +/-1 SE ribbon -> trend line -> mean
markers
Run ggsave(...) at the end to export --- or swap it for print(p) if working
interactively

# -- Load libraries --
library(ggplot2)
library(tidyr)
library(dplyr)

# -- 1. Enter the data --
raw <- data.frame(
  y0   = c(87.3, 59.0, 76.7, 70.6, 54.9, 78.2, 73.7, 61.8, 85.3, 82.3,
           68.6, 67.8, 66.2, 81.0, 72.3),
  y1   = c(86.9, 60.2, 76.5, 76.1, 55.1, 75.3, 70.8, 68.7, 84.4, 86.9,
           65.4, 69.2, 67.0, 82.3, 74.6),
  y2   = c(86.7, 60.0, 75.7, 72.1, 57.2, 69.1, 71.8, 68.2, 79.2, 79.4,
           72.3, 66.3, 67.0, 86.8, 75.3),
  y3   = c(75.5, 53.6, 69.5, 65.3, 49.0, 67.6, 74.6, 57.4, 67.0, 77.4,
           60.8, 57.9, 56.2, 73.9, 66.1),
  y0_1 = c(83.8, 65.3, 81.2, 75.4, 55.3, 70.3, 76.5, 66.0, 76.7, 77.2,
           67.3, 50.3, 57.7, 74.3, 74.0),
  y1_1 = c(85.5, 66.9, 79.5, 76.7, 58.3, 72.3, 79.9, 70.9, 79.0, 74.0,
           70.7, 51.4, 57.0, 77.7, 74.7),
  y2_1 = c(86.2, 67.0, 84.5, 74.3, 59.1, 70.6, 80.4, 70.3, 76.9, 77.8,
           68.9, 53.6, 57.5, 72.6, 74.5),
  y3_1 = c(81.2, 60.6, 75.2, 66.7, 54.2, 68.6, 71.6, 64.1, 70.3, 67.9,
           65.9, 48.0, 51.5, 68.0, 65.7)
)

# -- 2. Add a time index (row = observation point) --
raw$time <- seq_len(nrow(raw))

# -- 3. Pivot to long format --
long <- raw %>%
  pivot_longer(
    cols      = -time,
    names_to  = "variable",
    values_to = "value"
  ) %>%
  filter(!is.na(value)) %>%
  mutate(
    group     = if_else(grepl("_1$", variable), "Group 2", "Group 1"),
    timepoint = as.integer(sub(".*y(\\d).*", "\\1", variable))
  )

# -- 4. Compute per-variable summary (mean +/- SE across observations) --
summary_df <- long %>%
  group_by(group, timepoint) %>%
  summarise(
    mean_val = mean(value, na.rm = TRUE),
    se       = sd(value, na.rm = TRUE) / sqrt(n()),
    .groups  = "drop"
  )

# -- 5. Plot --
p <- ggplot(summary_df, aes(x = timepoint, y = mean_val,
                             color = group, fill = group)) +
  geom_ribbon(aes(ymin = mean_val - se, ymax = mean_val + se),
              alpha = 0.15, color = NA) +
  geom_line(linewidth = 1.1) +
  geom_jitter(data = long,
              aes(x = timepoint, y = value, color = group),
              width = 0.07, alpha = 0.25, size = 1.6, inherit.aes = FALSE) +
  geom_point(size = 3.5, shape = 21, color = "white", stroke = 1.5) +
  scale_x_continuous(breaks = 0:3,
                     labels = paste0("T", 0:3)) +
  scale_color_manual(values = c("Group 1" = "#2C7BB6", "Group 2" = "#D7191C")) +
  scale_fill_manual( values = c("Group 1" = "#2C7BB6", "Group 2" = "#D7191C")) +
  labs(
    title    = "Response Values Over Time",
    subtitle = "Mean +/- SE with individual observations",
    x        = "Time Point",
    y        = "Response Value",
    color    = NULL,
    fill     = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title       = element_text(face = "bold", size = 15),
    plot.subtitle    = element_text(color = "grey45", margin = margin(b = 10)),
    legend.position  = "top",
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "grey90")
  )

ggsave("plot.png", plot = p, width = 7, height = 5, dpi = 150)
cat("Plot saved to plot.png\n")
```
## Comparative Analysis

The difference between the plan-informed response and the generic response reveals
the mistakes that are possible if you neglect to create a plan for yourself or for
any generative artificial intelligence that is helping you to create code.

The first difference that I noticed between the two was Claude's 
misunderstanding of the data structure. Claude described the dataset as having "8
variables across 14 rows" and treated each row as a time point rather than as
an individual subject. This led to a some issues down the line once it created the code
the data was pivoted incorrectly.
The groups were labeled as "Group 1" and
"Group 2" rather than "Control" and "Treatment," no subject IDs were created,
and the visualization showed time on the x-axis as a numeric scale (0--3)
without labels. Claude entirely dropped the treatment groups 16th data because there was 
not a 16th women in the control group. 

The plan-informed response however created a useful code. It followed the steps
without much deviation and was able to make a better visualization that 
succesfully reached my initial goal which I provided to it at the beginning. 


# Reflection

Coming into this course I have had experience programming in the languages: 
Python, SQL, and some Java and HTML, but I had not had the opportunity to explore
R so this has been very exciting learning a new language throughout the course. 
Previously, I had not practiced writing out explicit plans involving goals, needs
or steps and I believe that this PCIP process will prove to be very beneficial as 
I continue working on code throughout my education and future career. 
Most of the previous courses that I've taken where we program we did not dive 
deep into many of the useful packages that the languages supported and being able
to get familiar with installing, loading and exploring packages has also been a
very useful experience. The data that we worked with for the homework assignments
have all been fun and interesting cases which made completing assignments very 
enjoyable. Overall, I have thoroughly enjoyed STAT 184, the lectures, content 
and assignments have all felt meaningful, useful, and kept me engaged through
the interesting data and fun activites. At the beginning of the course I had set
my own personal goal of working to the best of my ability to learn and understand
all material through giving my best efforts on all activities and readings and 
make a strong effort to earn the highest grade possible. I believe I have achieved
that goal, I am confident in my understanding of all of the material we have 
covered throughout the course and anticipate receiving a grade that I am pleased
with as a result of applying what I learned here into the assignments as well as
in my participation and success at the datafest event. I'd like to thank Dr. Bubb
and the teaching team for a great semester!

# Appendix: Code Appendix {#sec-code-appendix}

All code used to produce the analyses and visualizations in this document is below, organized by section.

## Busiest Airports: Data Scraping and Cleaning

```{r}
#| label: code-airports-scrape
#| eval: false

# ---- Load Required Packages ----
library(tidyverse)   # Data manipulation and visualization
library(rvest)       # Web scraping
library(janitor)     # Clean column names
library(knitr)       # Table formatting
library(kableExtra)  # Styled tables
library(ggplot2)     # Grammar of graphics

# ---- Step 1: Scrape Wikipedia Page for Airport Tables ----
# The Wikipedia page contains multiple tables; tables [[2]] through [[7]]
# correspond to the years 2025 through 2020.
airportRawList <- read_html(
  x = "https://en.wikipedia.org/wiki/List_of_busiest_airports_by_passenger_traffic"
) |>
  html_elements(css = "table") |>
  html_table()

# ---- Step 2: Extract Each Year's Table ----
# Apply clean_names() for consistent column naming and add a year identifier.
year_2025 <- airportRawList[[2]] |> clean_names() |> mutate(year = 2025)
year_2024 <- airportRawList[[3]] |> clean_names() |> mutate(year = 2024)
year_2023 <- airportRawList[[4]] |> clean_names() |> mutate(year = 2023)
year_2022 <- airportRawList[[5]] |> clean_names() |> mutate(year = 2022)
year_2021 <- airportRawList[[6]] |> clean_names() |> mutate(year = 2021)
year_2020 <- airportRawList[[7]] |> clean_names() |> mutate(year = 2020)

# ---- Step 3: Combine All Years into One Data Frame ----
all_airports <- bind_rows(year_2025, year_2024, year_2023,
                          year_2022, year_2021, year_2020)

# ---- Step 4: Clean and Filter ----
airports_clean <- all_airports |>
  mutate(
    # Remove reference annotations (e.g., "[1]") from passenger counts
    totalpassengers = str_remove_all(totalpassengers, "[^0-9]"),
    # Convert passenger counts from character to numeric
    totalpassengers = as.numeric(totalpassengers),
    # Replace NA rank changes with 0 (no change)
    rankchange = replace_na(rankchange, 0)
  ) |>
  # Select and reorder columns for clarity
  select(
    year, rank, code_iata_icao, airport, location,
    country, totalpassengers, rankchange
  ) |>
  # Retain only the six airports of interest
  filter(code_iata_icao %in% c(
    "ATL/KATL", "FRA/EDDF", "DXB/OMDB",
    "LAX/KLAX", "DFW/KDFW", "DEN/KDEN"
  ))
```

## Busiest Airports: Table Generation

```{r}
#| label: code-airports-table
#| eval: false

# ---- Build a Formatted Table ----
airports_clean |>
  arrange(year, rank) |>
  mutate(totalpassengers = format(totalpassengers, big.mark = ",")) |>
  kable(
    col.names = c(
      "Year", "Rank", "Code", "Airport", "Location",
      "Country", "Total Passengers", "Rank Change"
    ),
    align = "c"
  ) |>
  kable_styling(
    latex_options = c("striped", "hold_position", "scale_down"),
    full_width = FALSE
  )
```

## Busiest Airports: Plot Generation

```{r}
#| label: code-airports-plot
#| eval: false

# ---- Build the Scatter-Line Plot ----
ggplot(airports_clean, aes(
  x = totalpassengers,
  y = rank,
  color = code_iata_icao,
  group = code_iata_icao,
  shape = factor(year)
)) +
  geom_line(linewidth = 1) +          # Connect points by airport
  geom_point(size = 4) +              # Mark each year's observation
  scale_y_reverse() +                 # Rank 1 at the top
  scale_color_manual(
    values = c(
      "ATL/KATL"  = "#0072B2",
      "DFW/KDFW"  = "#D55E00",
      "DEN/KDEN"  = "#009E73",
      "LAX/KLAX"  = "#E69F00",
      "FRA/EDDF"  = "#CC79A7",
      "DXB/OMDB"  = "#56B4E9"
    ),
    labels = c(
      "ATL/KATL"  = "ATL",
      "DFW/KDFW"  = "DFW",
      "DEN/KDEN"  = "DEN",
      "LAX/KLAX"  = "LAX",
      "FRA/EDDF"  = "FRA",
      "DXB/OMDB"  = "DXB"
    )
  ) +
  scale_shape_manual(
    values = c(16, 17, 15, 18, 3, 7),
    name = "Year"
  ) +
  labs(
    x = "Total Passengers",
    y = "Global Rank",
    color = "Airport",
    shape = "Year"
  ) +
  theme_classic() +
  theme(
    text = element_text(size = 12),
    legend.position = "right",
    plot.title = element_text(face = "bold")
  )
```

## Monte Carlo Integration: Function Definition

```{r}
#| label: code-mc-function
#| eval: false

# ---- Define the Weibull PDF (shape = 1.5, scale = 1) ----
f <- function(x) {
  dweibull(x, shape = 1.5, scale = 1)
}

# ---- Monte Carlo Simulation Function ----
# Simulates n random points in the bounding rectangle [x_min, x_max] x [y_min, y_max]
# and estimates the integral of func over [x_min, x_max].
#
# Parameters:
#   n      - Number of random points to sample
#   func   - The function to integrate
#   x_min  - Lower bound of integration
#   x_max  - Upper bound of integration
#   y_min  - Lower bound of the sampling rectangle
#   y_max  - Upper bound of the sampling rectangle
#
# Returns: A list with $df (data frame of points) and $estimate (integral estimate)
monte_carlo_sim <- function(n, func, x_min, x_max, y_min, y_max) {
  # Generate uniform random points
  x_vals <- runif(n, min = x_min, max = x_max)
  y_vals <- runif(n, min = y_min, max = y_max)

  # Evaluate the function at each x
  f_x <- func(x_vals)

  # Classify each point as inside (under curve) or outside (above curve)
  df <- data.frame(
    x   = x_vals,
    y   = y_vals,
    f_x = f_x
  ) |>
    mutate(flag = ifelse(y <= f_x, "inside", "outside"))

  # Compute the integral estimate
  area       <- (x_max - x_min) * (y_max - y_min)
  proportion <- mean(y_vals <= f_x)
  estimate   <- area * proportion

  return(list(df = df, estimate = estimate))
}
```

## Monte Carlo Integration: Simulation and Visualization

```{r}
#| label: code-mc-viz
#| eval: false

# ---- Helper: Build a Single Monte Carlo Plot ----
build_mc_plot <- function(result, n_label) {
  ggplot(result$df, aes(x = x, y = y)) +
    geom_point(aes(color = flag), alpha = 0.4, size = 0.8) +
    stat_function(
      fun = dweibull,
      args = list(shape = 1.5, scale = 1),
      xlim = c(0, 4),
      color = "black",
      linewidth = 1
    ) +
    scale_color_manual(
      values = c("inside" = "#0072B2", "outside" = "#D55E00")
    ) +
    labs(
      x = "x",
      y = "f(x)",
      color = "Point Location",
      title = paste0("Monte Carlo Integration (", n_label, ")"),
      subtitle = paste("Estimate =", round(result$estimate, 4))
    ) +
    theme_minimal(base_size = 10) +
    theme(legend.position = "bottom")
}

# ---- Run Simulations at Four Resolutions ----
set.seed(42)  # For reproducibility

res_100   <- monte_carlo_sim(100,    f, 0, 4, 0, 0.8)
res_500   <- monte_carlo_sim(500,    f, 0, 4, 0, 0.8)
res_2000  <- monte_carlo_sim(2000,   f, 0, 4, 0, 0.8)
res_10000 <- monte_carlo_sim(10000,  f, 0, 4, 0, 0.8)

# ---- Display the Four-Pane Small Multiple ----
# In the rendered document, these four plots appear as a 2x2 grid
# using Quarto's layout-ncol and fig-subcap options.
build_mc_plot(res_100,   "n = 100")
build_mc_plot(res_500,   "n = 500")
build_mc_plot(res_2000,  "n = 2,000")
build_mc_plot(res_10000, "n = 10,000")
```
